Randomized trace estimation for parameter-dependent matrices applied to spectral density approximation

### 2025-02-14 paper/analysis.tex
push changes from Overleaf

\color{blue}
In this section, we consider a general parameter-dependent matrix
\begin{equation*}
    \mtx{B}(t) = \begin{bmatrix}
        b_{11}(t) & b_{12}(t) & \dots & b_{1n}(t) \\
        b_{21}(t) & b_{22}(t) & \dots & b_{2n}(t) \\
        \vdots & \vdots & \ddots & \vdots \\
        b_{n1}(t) & b_{n2}(t) & \dots & b_{nn}(t) \\
    \end{bmatrix} \in \mathbb{R}^{n \times n},
\end{equation*}
where each entry $b_{ij}(t)$ is a continuous function on a bounded interval $[a,b]$. \emph{If} these entries are given explicitly, we can simply compute the trace by its definition 
$\Trace(\mtx{B}(t)) = b_{11}(t)+ \cdots + b_{nn}(t)$. This becomes more difficult when $\mtx{B}(t)$ is given implicitly (e.g., as a matrix function) and can only be accessed through matrix-vector products. The common theme of all methods analyzed in this section is to extract trace estimates from multiplying $\mtx{B}(t)$ with fixed random vectors.

>>>>

\color{blue}
In this section, we consider a general parameter-dependent matrix
$\mtx{B}(t) \in \mathbb{R}^{n \times n}$, where each
entry $b_{ij}(t)$ is a 
%\begin{equation*}
%     \mtx{B}(t) = \begin{bmatrix}
%         b_{11}(t) & b_{12}(t) & \dots & b_{1n}(t) \\
%         b_{21}(t) & b_{22}(t) & \dots & b_{2n}(t) \\
%         \vdots & \vdots & \ddots & \vdots \\
%         b_{n1}(t) & b_{n2}(t) & \dots & b_{nn}(t) \\
%     \end{bmatrix} \in \mathbb{R}^{n \times n},
% \end{equation*}
continuous function on a bounded interval $[a,b]$. ??? UNTIL HERE ??? \emph{If} these entries are given explicitly, we can simply compute the trace by its definition 
$\Trace(\mtx{B}(t)) = b_{11}(t)+ \cdots + b_{nn}(t)$. This becomes more difficult when $\mtx{B}(t)$ is given implicitly (e.g., as a matrix function) and can only be accessed through matrix-vector products. The common theme of all methods analyzed in this section is to extract trace estimates from multiplying $\mtx{B}(t)$ with fixed random vectors.
